---
title: 'BricolR : T=Rasters'
author: "Philémon Poux"
date: "28/05/2019"
output: html_document
---

```{r setup, include=FALSE}


load.lib<-c("raster","rgdal","tidyverse","gdalUtils","RColorBrewer","wesanderson","sf","stars")


install.lib<-load.lib[!load.lib %in% installed.packages()]
for(lib in install.lib) install.packages(lib,dependencies=TRUE)
sapply(load.lib,require,character=TRUE)




```

## The basic of rasters with R
  We are going to import population datasets for France and Spain and explore these rasters a little bit
```{r}
france__f_65plus <- raster("france_f_65plus.tif")
france_m_65plus <- raster("france_m_65plus.tif")
esp_65plus <- raster("spain_65plus.tif")

###Do both file for France have the same extent, resolution etc ?
extent(france__f_65plus)
extent(france_m_65plus)

###Plotting the rasters, choose your favorite palette and plot the results. You may want to break results with breaks
plot(france__f_65plus)
plot(france_m_65plus)

###Do the same thing with ggplot
france_df <- as.data.frame(france__f_65plus,xy=T) 
ggplot(france_df)+geom_raster(aes(x=x,y=y,fill=france_f_65plus))

###What is the max number of women older than 65 in France per cell ?
cellStats(france__f_65plus,max)


###Access the dataframe of all the values with the getValues() function 
values <- getValues(france__f_65plus)





```

## Raster Operations with R
We are now going to manipulate the rasters in order to compute the sum of men and women, have a map with Spain and France, crop the data to one region

```{r}
###Create a raster with the sum of men and women above 65 in France. Careful, you need to make sure both files are comparable to do so.
res(france__f_65plus)
res(france_m_65plus)
origin(france__f_65plus)
origin(france_m_65plus)

france_m_65plus_agg <- aggregate(france_m_65plus,fact=2.5,fun=sum,na.rm=T)
res(france_m_65plus_agg)

france_m_65plus_res <- resample(france_m_65plus,france__f_65plus)
res(france_m_65plus_res)
compareRaster(france__f_65plus,france_m_65plus_res)
france_65plus <- france__f_65plus+france_m_65plus_res
plot(france_65plus)

###Merge the data, careful, you might need to adjust the origins
plot(esp_65plus)
res(esp_65plus)

origin(esp_65plus) <- origin(france_65plus)

europe_65plus <- merge(france_65plus,esp_65plus)
plot(europe_65plus,main="Population over 65 in France and Spain")

###See the data only in Île-de-France
extent_idf <- extent(c(1.44728,3.555687,48.12054,49.23719))
plot(crop(france_m_65plus,extent_idf))

```
## Rasters and Vectors
Let's see how you can manipulate both rasters and vectors data. We will mask a raster within a region, extract the values depending the regions, transform a raster into vector data...


```{r cars}
#Import the shapefile of the regions and check that it has the same CRS as our data
regions <- st_read("regions/regions.shp")
plot(regions)

###Isolate the Occitanie and mask the data within this region
Occitanie <- regions[which(regions$nom=="Occitanie"),]
plot(crop(mask(europe_65plus,Occitanie),Occitanie))

###Create contours for the 5 quantiles for europe, create a SpatialDataPoints of all cells that have a value 
regionsSpatials <- as(regions,"Spatial")

raster_regions <- rasterize(regionsSpatials,france_m_65plus)
plot(raster_regions)

breaks <- quantile(france__f_65plus)
contours <- rasterToContour(france__f_65plus,levels=breaks)
plot(france__f_65plus)
plot(contours,add=T)

points <- rasterToPoints(france__f_65plus,fun=function(x) x>10000)
plot(points)

###Extract the sum of population per regions in France
extraction <- raster::extract(france_65plus,regions,fun=sum,na.rm=T)
extraction
regions$sumpop <- extraction
ggplot(regions)+geom_sf(aes(fill=sumpop))

```

## Saving your work
You can save your raster, stack different rasters with a stack or just save your working environment to come back to it later
```{r pressure}
###save your file
writeRaster(x = europe_65plus ,filename = "europe_65plus.tif",overwrite=T)

###create a stack with the data for women, men and the sum. Then save it into a multi layer file

###Save your working environment for later, you can access it with load
save.image("rasters.RData")

```

## GDAL
For those willing to see how gdal works, I propose that we explore it together

```{r}
file_fra_large <- "C:/Users/Ameline/Desktop/Philemon/Databases/WorldPop/FRA/fra_f_65_80_2010.tif"
file_spa_large <- "C:/Users/Ameline/Desktop/Philemon/Databases/WorldPop/ESP/esp_f_65_80_2010.tif"

list_large <- c(file_fra_large,file_spa_large)
france_large <- raster(file_fra_large)
spain_large <- raster(file_spa_large)
france_large <- raster(file_fra_large)
origin(spain_large)
origin(france_large)
origin(spain_large) <- origin(france_large)
system.time(merge(france_large,spain_large))

merge_gdal <- function(lst) {
  gdalbuildvrt(lst,"temp.vrt")
  mosaic_rasters("temp.vrt","mosaic_gdal.tif",verbose=T,co=c("COMPRESS=LZW","BITS=12","TILED=YES"), ot="Float32",overwrite=T)
  file.remove("temp.vrt")
}
system.time(merge_gdal(list_large))

crs(france_large)
new_crs <- CRS("+proj=laea +x_0=0 +y_0=0 +lon_0=0 +lat_0=0")
system.time(truc <- projectRaster(france_large,crs=new_crs))
system.time(gdalwarp(srcfile = file_fra_large,dstfile = "france_reprojected.tif",t_srs = new_crs,verbose=T,overwrite = T))


```


